---
title: "Practical Machine Learning Assignment"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
```

### Coursera Data Science Specialization
### Practical Machining Learning Module
### Project Assignment

#### 1. Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 

#### 2. Data
The training data for this project are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: 
http://groupware.les.inf.puc-rio.br/har

#### 3. Load required packages for analysis
```{r load packages}
library(caret)
library(rpart)
library(ggplot2)
library(corrplot)
library(randomForest)
library(rattle)
set.seed(12345)
```

#### 4. Load data
```{r load data}
training_raw <- read.csv("pml-training.csv")[,-1]
testing <- read.csv("pml-testing.csv")[,-1]
# check dimension of the training and test dataset
dim(training_raw)
dim(testing)
```

#### 5. Clean data
```{r cleaning data}
# remove predictors that have many missing/NA values or non-unique values
NZV <- nearZeroVar(training_raw)
training <- training_raw[, -NZV]
testing <- testing[, -NZV]

# remove cases that have many missing/NA values
NaValues <- sapply(training, function(x) mean(is.na(x))) > 0.9
training <- training[, NaValues == "FALSE"]
testing <- testing[, NaValues == "FALSE"]

# remove id and time variables
training <- training[,-c(1:5)]
testing <- testing[,-c(1:5)]

# check dimension of the cleaned up dataset
dim(training)
dim(testing)

# take a look at the training dataset
head(training)
```

#### 6. Prepare data partition, for later validation
```{r prepare cross validation dataset}
inTrain <- createDataPartition(y= training$classe, p = 0.7, list = FALSE)
training <- training[inTrain, ]
crossvalidation <- training[-inTrain, ]
```

#### 7. Now we can train our models given the preprocess with PCA
```{r Train models}
# decision trees
model_tree <- train(classe~., data = training, method = "rpart")
# print result of model prediction on original training and crossvalidation dataset
predict_training_tree <- predict(model_tree, training)
confusionmatrix_training_tree <- confusionMatrix(predict_training_tree, training$classe)

predict_crossvalidation_tree <- predict(model_tree, crossvalidation)
confusionmatrix_cv_tree <- confusionMatrix(predict_crossvalidation_tree, crossvalidation$classe)

print(confusionmatrix_cv_tree)

# random forest
model_rf <- train(classe~., data = training, method = "rf")
# print result of model prediction on original training and crossvalidation dataset
predict_training_rf <- predict(model_rf, training)
confusionmatrix_training_rf <- confusionMatrix(predict_training_rf, training$classe)

predict_crossvalidation_rf <- predict(model_rf, crossvalidation)
confusionmatrix_cv_rf <- confusionMatrix(predict_crossvalidation_rf, crossvalidation$classe)

print(confusionmatrix_cv_rf)
```

#### 8. Conclusion
The confusionmatrix showed that the accuracy of the random forest models is better than the decision tree model. 
Therefore, we used this model to predict on the testing dataset. 

#### 9. Predict on testing dataset
```{r predict on testing dataset}
predict_testing <- predict(model_rf, testing)
predict_testing
```

#### 10. Appendix
``` {r Principal component analysis}
# explore the remianing predictors
# check the factor variables
predictor_factor <- which(sapply(training, class) == "factor")
# explore correlation between predictors
predictor_cor <- abs(cor(training[,-predictor_factor]))
# turn lower tri to 0
predictor_cor[lower.tri(predictor_cor, diag = TRUE)] <- 0
# visualize result
corrplot(predictor_cor, method  = "color", type = "upper", cl.lim = c(0,1), tl.col = rgb(0, 0, 0))
# find highly correlated predictors
which(predictor_cor > 0.8, arr.ind = TRUE)
# Therefore, there are highly correlated predictors, principal component analysis is necessary.
```